---
title: "Forecasting Lyme's Disease"
author: "Jeremy Ling and Alexis Rivera-Andrade"
date: "May 9, 2018"
output:
  slidy_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is Lyme disease?
Lyme disease is caused by bacteria and transmitted by infected black legged and deer ticks.\newline
Symptoms usually occur from 3 to 30 days after the bite.
Symptoms include:

               -Severe headaches

                -Rashes all over body. Erythema Migrans

                -Arthritis

                -Irregular heartbeat

                -Inflammation of the brain and spinal cord

                -Shooting pains, numbness, or tingling in the feet

## 14 Top states
-95 percent of the confirmed Lyme diseases cases were reported from these 14 states

    Connecticut
    Delaware
    Maine
    Maryland
    Massachusetts
    Minnesota
    New Hampshire
    New Jersey
    New York
    Pennsylvania
    Rhode Island
    Vermont
    Virginia
    Wisconsin

-Only Wisconsin and Rhode Island have monthly data reported for Lyme disease.

- We gathered the data from the Wisconsin Department of Health Services

## Google Trends

-We use google trends on search topics "Lyme's Disease" in Wisconsin from 2009 to 2016

-Limitation: The data is normalized

```{r cars, echo = TRUE}
library(rmarkdown)
library(forecast)
library(stats)
library(stargazer)
library(vars)
library(shiny)
library(ggplot2)
library(tseries)
#### Lyme disease Data ####
WI09 <- c(16, 23, 26, 36, 115, 464, 647, 286, 148, 82, 60, 45)
WI10 <- c(39, 30, 38, 72, 233, 821, 699, 239, 131, 133, 44, 32)
WI11 <- c(39, 21, 38, 67, 152, 589, 832, 346, 156, 88, 63, 34)
WI12 <- c(41, 33, 57, 93, 200, 453, 341, 108, 83, 38, 25, 15)
WI13 <- c(20, 11, 14, 13, 109, 398, 594, 284, 167, 71, 28, 17)
WI14 <- c(16, 9, 15, 27, 85, 254, 320, 115, 78, 45, 11, 9)
WI15 <- c(13, 8, 11, 33, 144, 376, 373, 142, 77, 74, 28, 22)
WI16 <- c(9, 14, 28, 57, 127, 433, 438, 146, 80, 91, 47, 21)

WI <- c(WI09, WI10, WI11, WI12, WI13, WI14, WI15, WI16)

WI2 <- WI / 8.32

# Year and Month
Year <- rep(x = c(2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016),
            each = 12)
Month <- rep(x = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12),
             times = 8)

#### Temperature Data ####
origin <- read.csv("trend_data.csv", skip = 2, header = TRUE)
trend <- origin[61:156,]

#### Master List ####
# master set
master <- data.frame(cbind(trend, WI, WI2))
names(master) = c("Trend", "Reports", "W.Reports")

# declaring time series data
lyme <- ts(master$W.Reports,
           start = c(2009, 1),
           frequency = 12)

master <- ts(master[-2],
             start = c(2009, 1),
             frequency = 12)
```
## Initial Plot

```{r pressure}

# initial plots
plot(lyme)
lines(trend)
```

## Linear
```{r}
# training data
lyme.a<-window(lyme, start = c(2009,1), end = c(2014,12))

# linear regression
lyme.lm <- tslm(formula = lyme.a ~ trend)
print(summary(lyme.lm))

lyme.f1 <- forecast(lyme.lm, h = 24, level = 95)
lyme.prd1 <- lyme.lm$fit

plot(lyme.f1)
lines(lyme.prd1)
lines(lyme)

```

##Quadratic
```{r}
lyme.qt <- tslm(lyme.a~trend+I(trend*trend))
print(summary(lyme.qt))

lyme.f2<- forecast(lyme.qt, h = 24, level = 95)
lyme.prd2<-lyme.qt$fit

plot(lyme.f2)
lines(lyme.prd2)
lines(lyme)

```

##LogLinear
```{r}
loglyme <- log(lyme)
loglyme.a<-window(loglyme, start = c(2009,1), end = c(2014,12))

lyme.log <- tslm(loglyme.a~trend)
print(summary(lyme.log))

lyme.f3 <-forecast(lyme.log, h = 24,level = 95)

lyme.prd3 <- lyme.log$fit
plot(lyme.f3)
lines(loglyme)
lines(lyme.prd3)
```


##AIC and BIC scores
```{r}
data.frame(AIC = c(AIC(lyme.lm),AIC(lyme.qt), AIC(lyme.log)), 
           BIC = c(BIC(lyme.lm), BIC(lyme.qt), BIC(lyme.log)), row.names = 
             c("Linear", "Quadratic", "Log"))
```

##Seasonality
```{r}
#### Seasonality ####

# linear regression
lyme.lm <- tslm(formula = lyme.a  ~ season)
summary(lyme.lm)

# fitted values
lyme.prd <- lyme.lm$fit

# forecast
lyme.f <- forecast(lyme.lm, h=24, level = 95)

# plotting data
plot(lyme.f,col = "blue", lwd = 2)
lines(lyme, lwd = 2)
lines(lyme.prd, col = "red", lwd = 2)

```



##Stationality 
Adf Test:
```{r}
adf.test(x = lyme)
```
ndiffs
```{r}
ndiffs(lyme)
```
##ACF and PACF
```{r}
# ACF and PACF
acf(lyme)
pacf(lyme)
```

##Fitting Model
```{r}
fit <- auto.arima(lyme)
res <- residuals(fit)

lyme.f2 <- forecast(fit, h = 24, level = 95)
plot(lyme.f2, col = "blue", lwd = "2")

```

